---
title: "Homework_05"
author: "Andrew McCracken"
date: "2/16/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

[Home Page](index.html)

# Working with Matrices, Lists, and Data Frames

### 1. Assign to the variable n_dims a single random integer between 3 and 10.
```{r}
# floor rounds down to nearest integer, so setting from 3 to 11 will return a 10 if it is 10.999
n_dims <- floor(runif(1,3,11))
n_dims
```
* Create a vector of consecutive integers from 1 to n_dims2
```{r}
myvec <- seq(1,(n_dims)^2)
myvec
```
* Use the sample function to randomly reshuffle these values.
```{r}
myvec <- sample(myvec)
myvec
```
* create a square matrix with these elements.
```{r}
m <- matrix(myvec, nrow=sqrt(length(myvec)))
m
```
* find a function in r to transpose the matrix.
```{r}
m <- t(m)
m
```
* calculate the sum and the mean of the elements in the first row and the last row.
```{r}
sum(m[1,])
sum(m[-1,])
mean(m[1,])
mean(m[-1,])
```
* read about the eigen() function and use it on your matrix  
```{r}
eigen_m <- eigen(m)
```
* look carefully at the elements of `$values` and `$vectors`. What kind of numbers are these?  
    * `$values` is the mathematical variance of a metrix and how symmetrical the matrix is. 
    * `$vectors` is a special set of scalars associated with a linear system of equations. a vector whose direction remains unchanged when a linear transformation is applied to it.
    
* dig in with the typeof() function to figure out their type.
```{r}
typeof(eigen_m$values)
typeof(eigen_m$vectors)
# typeof returns "complex" for each of these values which are also labled as 'doubles'
```
* if have set your code up properly, you should be able to re-run it and create a matrix of different size because n_dims will change.


### 2. Create a list with the following named elements:

* mymatrix, which is a 4 x 4 matrix filled with random uniform values
* mylogical which is a 100-element vector of TRUE or FALSE values. Do this efficiently by setting up a vector of random values and then applying an inequality to it.
* my_letters, which is a 26-element vector of all the lower-case letters in random order.
```{r}
mylist <- list(mymatrix = matrix(data = runif(16), ncol = 4), 
            mylogical = c(runif(100) < 0.5),
            myletters = letters[1:26])
```
* create a new list, which has the element[2,2] from the matrix, the second element of the logical vector, and the second element of the letters vector.
```{r}
newlist <- list(mylist$mymatrix[2,2],
                mylist$mylogical[2],
                mylist$myletters[2])
newlist
```
* use the typeof() function to confirm the underlying data types of each component in this list
```{r}
typeof(newlist[[1]])
typeof(newlist[[2]])
typeof(newlist[[3]])
```
* combine the underlying elements from the new list into a single atomic vector with the c() function.
```{r}
newvec <- c(unlist(newlist))
newvec
```
* what is the data type of this vector?
```{r}
typeof(newvec)
```

### 3. Create a data frame with two variables (= columns) and 26 cases (= rows).
* call the first variable my_unis and fill it with 26 random uniform values from 0 to 10  
* call the second variable my_letters and fill it with 26 capital letters in random order.
```{r}
myunits <- runif(26,0, 10)
my_letters <- sample(LETTERS[1:26]) 
mydata <- data.frame(myunits, my_letters)
mydata
```

* for the first variable, use a single line of code in R to select 4 random rows and replace the numerical values in those rows with NA.

```{r}
mydata$myunits[sample(1:26, 4)] <- NA
```

* for the first variable, write a single line of R code to identify which rows have the missing values.

```{r}
which(is.na(mydata$myunits))
```

* for the second variable, sort it in alphabetical order
```{r}
mydata <- mydata[order(mydata$my_letters),]
mydata
```

* calculate the column mean for the first variable.
```{r}
summary(mydata$myunits)
#or
mean(mydata$myunits, na.rm = TRUE)
```














