---
title: "Homework_06"
author: "Andrew McCracken"
date: "2/23/2022"
output: html_document
---
[Home Page](index.html)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Read in data
```{r}
library(ggplot2) # for graphics
library(MASS) # for maximum likelihood estimation

starfish <- read.table("C:/Users/andre/Desktop/Comp-Bio_381/TableS3_FigS3_data.csv",header=TRUE,sep=",")
str(starfish)
summary(starfish)
```
  
### Plotting Histogram Data

```{r}
p1 <- ggplot(data=starfish, aes(x=arm_length, y=..density..)) +
  geom_histogram(color="grey60",fill="cornsilk",size=0.2) 
print(p1)
```

### Add empirical density curve

```{r}
p1 <-  p1 +  geom_density(linetype="dotted",size=0.75)
print(p1)  
```

### Get maximum likelihood parameters for normal

```{r}
normPars <- fitdistr(starfish$arm_length,"normal")
print(normPars)
str(normPars)
normPars$estimate["mean"] # note structure of getting a named attribute
normPars$estimate["sd"]
```

### Plot `normal` probability density

```{r}
meanML <- normPars$estimate["mean"]
sdML <- normPars$estimate["sd"]

xval <- seq(0,max(starfish$arm_length),len=length(starfish$arm_length))

 stat <- stat_function(aes(x = xval, y = ..y..), fun = dnorm, colour="red", n = length(starfish$arm_length), args = list(mean = meanML, sd = sdML))
 p1 + stat
```

### Plot `exponential` probability density

```{r}
expoPars <- fitdistr(starfish$arm_length,"exponential")
rateML <- expoPars$estimate["rate"]

stat2 <- stat_function(aes(x = xval, y = ..y..), fun = dexp, colour="blue", n = length(starfish$arm_length), args = list(rate=rateML))
 p1 + stat + stat2
```

### Plot `uniform` probability density

```{r}
stat3 <- stat_function(aes(x = xval, y = ..y..), fun = dunif, colour="darkgreen", n = length(starfish$arm_length), args = list(min=min(starfish$arm_length), max=max(starfish$arm_length)))
 p1 + stat + stat2 + stat3
```

### Plot `gamma` probability density

```{r}
gammaPars <- fitdistr(starfish$arm_length,"gamma")
shapeML <- gammaPars$estimate["shape"]
rateML <- gammaPars$estimate["rate"]

stat4 <- stat_function(aes(x = xval, y = ..y..), fun = dgamma, colour="brown", n = length(starfish$arm_length), args = list(shape=shapeML, rate=rateML))
 p1 + stat + stat2 + stat3 + stat4
```

### Plot `beta` probability density

```{r}
pSpecial <- ggplot(data=starfish, aes(x=arm_length/(max(arm_length + 0.1)), y=..density..)) +
  geom_histogram(color="grey60",fill="cornsilk",size=0.2) + 
  xlim(c(0,1)) +
  geom_density(size=0.75,linetype="dotted")

betaPars <- fitdistr(x=starfish$arm_length/max(starfish$arm_length + 0.1),start=list(shape1=1,shape2=2),"beta")
shape1ML <- betaPars$estimate["shape1"]
shape2ML <- betaPars$estimate["shape2"]

statSpecial <- stat_function(aes(x = xval, y = ..y..), fun = dbeta, colour="orchid", n = length(starfish$arm_length), args = list(shape1=shape1ML,shape2=shape2ML))
pSpecial + statSpecial
```

### What distribution best fits this dataset?: 

The best fitting distribution for this dataset is the gamma distribution, although the normal distribution is close as well. 

### Simulate a new dataset. 

```{r}
newdata <- rgamma(n=(length(starfish$arm_length)),shape=shapeML, rate=rateML)
newdata <- data.frame(1:(length(starfish$arm_length)),newdata)
names(newdata) <- list("ID","myVar")
newdata <- newdata[newdata$myVar>0,]
str(newdata)
summary(newdata$myVar)

p2 <- ggplot(data=newdata, aes(x=myVar, y=..density..)) +
  geom_histogram(color="grey60",fill="cornsilk",size=0.2) 
print(p2)

```


